---
title: "ch13_Geomarketing"
author: "Susan Paykin"
date: "6/2/2020"
output: pdf_document
---


## 13.1 Introduction

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(sf)
library(dplyr)
library(purrr)
library(raster)
library(osmdata)
library(spDataLarge)
```

- Geomarketing, sometimes referred to as location analysis or location intelligence
  - Example: where to locate a new shop
- People are fundamental to location analysis, but can also be applied to ecological concepts and models

## 13.2 Case Study: Bike Shops in Germany

Starting a chain of bike shops in Germany: where should the stores be opened, in order to attract any many customers as possible? Target audience: single young males (aged 20-40)


## 13.3 Tidy Input Data

```{r}
download.file("https://tinyurl.com/ybtpkwxz", 
              destfile = "census.zip", mode = "wb")
unzip("census.zip") # unzip the files
census_de = readr::read_csv2(list.files(pattern = "Gitter.csv"))

# pop = population, hh_size = household size
input = dplyr::select(census_de, x = x_mp_1km, y = y_mp_1km, pop = Einwohner,
                      women = Frauen_A, mean_age = Alter_D,
                      hh_size = HHGroesse_D)
# set -1 and -9 to NA
input_tidy = mutate_all(input, list(~ifelse(. %in% c(-1, -9), NA, .)))

```

## Create Census Rasters

```{r}
# create raster brick
input_ras = rasterFromXYZ(input_tidy, crs = st_crs(3035)$proj4string)

# simple plot
plot(input_ras)

# view raster details
input_ras

# reclassify values of rasters in accordance with the survey
rcl_pop = matrix(c(1, 1, 127, 2, 2, 375, 3, 3, 1250, 
                   4, 4, 3000, 5, 5, 6000, 6, 6, 8000), 
                 ncol = 3, byrow = TRUE)
rcl_women = matrix(c(1, 1, 3, 2, 2, 2, 3, 3, 1, 4, 5, 0), 
                   ncol = 3, byrow = TRUE)
rcl_age = matrix(c(1, 1, 3, 2, 2, 0, 3, 5, 0),
                 ncol = 3, byrow = TRUE)
rcl_hh = rcl_women
rcl = list(rcl_pop, rcl_women, rcl_age, rcl_hh)

# for loop to apply the reclassification matrix to the corresponding raster layer
reclass = input_ras
for (i in seq_len(nlayers(reclass))) {
  reclass[[i]] = reclassify(x = reclass[[i]], rcl = rcl[[i]], right = NA)
}

# ensure reclass layers have the same name as layers of input_ras
names(reclass) = c("pop", "women", "mean_age", "hh_size")

# view
reclass

```

## 13.5 Define Metropolitan Areas

- Define metropolitan areas as pixels of 20 sq. km inhabited by more than 500,000 people
- Pixels at this coarse resolution can be created using `aggregate()`
- Use argument `fact = 20` to reduce resolution of the result twenty-fold (note: the original raster resolution was 1 sq. km)

```{r}

# aggregate pixels, reduce resolution by 20x
pop_agg = aggregate(reclass$pop, fact = 20, fun = sum)

# keep only cells with > half million people
summary(pop_agg)

pop_agg = pop_agg[pop_agg > 500000, drop = FALSE]

# plot 
polys = pop_agg %>% 
  clump() %>%
  rasterToPolygons() %>%
  st_as_sf()

plot(polys)

metros = polys %>%
  group_by(clumps) %>%
  summarize()

plot(metros)

metros_wgs = st_transform(metros, 4326)

coords = st_centroid(metros_wgs) %>%
  st_coordinates() %>%
  round(4)

library(revgeo)
metro_names = revgeo(longitude = coords[, 1], latitude = coords[, 2], 
                     output = "frame")

metro_names = dplyr::pull(metro_names, city) %>% 
  as.character() %>% 
  ifelse(. == "Wülfrath", "Duesseldorf", .)


```


## 13.6 Points of Interest

```{r}

# load data
shops <- data("shops", package = "spDataLarge")

shops = map(metro_names, function(x) {
  message("Downloading shops of: ", x, "\n")
  # give the server a bit time
  Sys.sleep(sample(seq(5, 10, 0.1), 1))
  query = opq(x) %>%
    add_osm_feature(key = "shop")
  points = osmdata_sf(query)
  # request the same data again if nothing has been downloaded
  iter = 2
  while (nrow(points$osm_points) == 0 & iter > 0) {
    points = osmdata_sf(query)
    iter = iter - 1
  }
  points = st_set_crs(points$osm_points, 4326)
})

# checking if we have downloaded shops for each metropolitan area
ind = map(shops, nrow) == 0
if (any(ind)) {
  message("There are/is still (a) metropolitan area/s without any features:\n",
          paste(metro_names[ind], collapse = ", "), "\nPlease fix it!")}

# select only specific columns
shops = map(shops, dplyr::select, osm_id, shop)
# putting all list elements into a single data frame
shops = do.call(rbind, shops)

shops = st_transform(shops, proj4string(reclass))

# create poi raster
poi = rasterize(x = shops, y = reclass, field = "osm_id", fun = "count")

# construct reclassification matrix
int = classInt::classIntervals(values(poi), n = 4, style = "fisher")
int = round(int$brks)
rcl_poi = matrix(c(int[1], rep(int[-c(1, length(int))], each = 2), 
                   int[length(int)] + 1), ncol = 2, byrow = TRUE)
rcl_poi = cbind(rcl_poi, 0:3)  
# reclassify
poi = reclassify(poi, rcl = rcl_poi, right = NA) 
names(poi) = "poi"

```

## 13.7 Identifying Suitable Locations

- Add `poi` to the `reclass` raster stack and remove the population layer from it

```{r}

# add poi raster
reclass = addLayer(reclass, poi)
# delete population raster
reclass = dropLayer(reclass, "pop")

# calculating the final score!
result = sum(reclass)

result

```

## Exercises

```{r}

library(sp)

SPinput_tidy <- 
  input_tidy %>% 
  st_as_sf(coords = c("x", "y"), 
           crs = 3035) %>% 
  as("Spatial")

gridded(SPinput_tidy) <- TRUE

SPinput_brick <- brick(SPinput_tidy)


```

